Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-01 article-models-tt.tex
starting to include the extension subsection to pretty, cool new commands

The identity function of an extended family is the extension of the identity
functions on the components in the apropriate way. To find out what the
apropriate way is, note that
\begin{align*}
\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
\end{align*}
We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
\begin{align*}
\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
& \jdeq \ctxwk{P}{P},
\end{align*}
where we find the term $\idfunc[P]$. Therefore, we have the following inference rule:
\begin{equation}\label{idfunc-ext-comp}
\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}}
\end{equation}

@@

There is also a decomposition of terms $\jterm{\Gamma}{\ctxext{A}{P}}{w}$ into
components: the projections.

@@

@@

Pretty type theory is pretty as an ugly little duckling. At first, it is actually uglier than
structural type theory because one has to keep track of a variable nobody is
actually interested in. The situation improves when we introduce the operators
of extension and weakening, because from that point onwards one can draw notational
advantages from having the variable around.

@@

>>>>

We won't state a compatibility rule stating that the identity function is
compatible with extension because we will be able to prove that. Instead, we
will just state the compatibility rules for the identity function combined with
weakening and with substitution.

@@

We have seen above that the pairing function into $\ctxext{A}{P}$ is just the identity function on
$\ctxext{A}{P}$. To analyze the pairing functin a little further, we will also
need the projection maps from $\ctxext{A}{P}$ to $A$ and from $\ctxext{A}{P}$
to $P$. We will now define these and see that the identity function of an
extended family is the extension (or pairing) of the identity
functions on the components in the apropriate way.

To find out what the
apropriate way is, note that
\begin{align*}
\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}} & \jdeq \ctxwk{P}{{A}{\ctxext{A}{P}}}\\
& \jdeq \ctxext{\ctxwk{P}{{A}{A}}}{\ctxwk{P}{{A}{P}}}
\end{align*}
We have the term $\jterm{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{P}{A}}{\ctxwk{P}{\idfunc[A]}}$.
Thus we need to find out what $\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}}$ is:
\begin{align*}
\subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{P}}} & \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{P}}}\\
& \jdeq \ctxwk{P}{P},
\end{align*}
where we find the term $\idfunc[P]$. Therefore we define:

@@

We find the following inference rule, which expresses that the identity function
is compatible with extension:

\begin{lem}
For any $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$ we have
\begin{equation}\label{idfunc-ext-comp}
\inference{\jfam{\Gamma}{A}\quad\jfam{\ctxext{\Gamma}{A}}{P}}{\jtermeq{\ctxext{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}}{\idfunc[\ctxext{A}{P}]}{\subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}}}
\end{equation}
\end{lem}

\begin{proof}
Consider the families $\jfam{\Gamma}{A}$ and $\jfam{\ctxext{\Gamma}{A}}{P}$. Then
we have the judgmental equalities
\begin{align*}
\subst{\cprojsndf{A}{P}}{{\cprojfstf{A}{P}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}
& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\idfunc[\ctxwk{\ctxext{A}{P}}{\ctxext{A}{P}}]}}\\
& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{\idfunc[\ctxext{A}{P}]}}}\\
& \jdeq \subst{\idfunc[P]}{{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{\idfunc[\ctxext{A}{P}]}}}}\\
& \jdeq \subst{\idfunc[P]}{\ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}}}\\
& \jdeq \subst{\idfunc[A]}{\ctxwk{A}{\idfunc[\ctxext{A}{P}]}}\\
& \jdeq \idfunc[\ctxext{A}{P}].
\end{align*}
\end{proof}

@@

Pretty type theory is pretty as the ugly little duckling that grew up. At first, it is actually uglier than
structural type theory because one has to keep track of a variable nobody is
actually interested in. The situation improves when we introduce the operators
of extension and weakening, because from that point onwards one can draw notational
advantages from having the variable around.

@@

\begin{align}
& \inference{\jvfam{i}{\Gamma}{A}}{\jvfamcombi{{i}{x}}{{\Gamma}{A}}{P}}
& & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jfameq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}{A}}{\ctxext{\Delta}{B}}}\\
& \inference{\jfam{\ctxext{\Gamma}{A}}{P}}{\jfam{\Gamma}{\ctxext{A}{P}}}
& & \inference{\jfameq{\Gamma}{A}{B}\qquad\jfameq{\ctxext{\Gamma}{A}}{P}{Q}}{\jfameq{\Gamma}{\ctxext{A}{P}}{\ctxext{B}{Q}}}
\end{align}
